---
title: "Chapter-3 Exercise-1"
author: "Danish Salam"
date: "July 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling- Exercise#1

This document contains the tasks which are requsted in execrise#1 in chapter#3.

### Load the data in RStudio
Save the data set as a CSV file called refine_original.csv and load it in RStudio into a data frame.

```{r message = FALSE}
#load library
library(dplyr)
library(tidyr)

#load refine csv and hold it in local data frame.
org_df = read.csv("refine_original.csv")
df <- tbl_df(org_df)
```


### Task # 1: Clean up brand names

Clean up the 'company' column so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).

```{r df}
df$company <-gsub("\\w+ps$", "philips", df$company, ignore.case = TRUE)
df$company <-gsub("^ak\\s*\\w+", "akzo", df$company, ignore.case = TRUE)
df$company <-gsub("^van\\s\\w+", "van houten", df$company, ignore.case = TRUE)
df$company <-gsub("^un\\w+", "unilever", df$company, ignore.case = TRUE)

#unique company name list
df %>% select(company) %>% unique()
```

### Task#2: Separate product code and number

Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively

```{r}
#split the code number column into two columns prodcut code and number respectively.
df <- df %>%
  separate(Product.code...number, c("ProductCode", "ProductNumber"), "-")
df %>% select(ProductCode, ProductNumber) %>% unique()
```

### Task#3: Add product categories

```{r}
#add a new column product category. Product category will be identified by this product code.
df <- df %>% 
  mutate(category = case_when(
    .$ProductCode=="p" ~ "Smartphone",
    .$ProductCode=="v" ~ "TV",
    .$ProductCode=="x" ~ "Laptop",
    .$ProductCode=="q" ~ "Tablet"
  ))

#unique prodcut category list
df %>% select(category) %>% unique()

```

### Task#4: Add full address for geocoding

```{r}

df <-df %>% 
  mutate(full_address = paste(address, city, country, sep = ', '))
df %>% 
  select(full_address) %>% distinct
```

### Task#5: Create dummy variables for company and product category

```{r}
df <- df %>% 
  mutate(company_philips = if_else(company=="philips", 1, 0)) %>% 
  mutate(company_akzo = if_else(company=="akzo", 1, 0)) %>% 
  mutate(company_van_houten = if_else(company=="van houten", 1, 0)) %>% 
  mutate(company_unilever = if_else(company=="unilever", 1, 0)) %>% 
  mutate_at(vars(starts_with("company_")),funs(as.logical)) %>%
  select(company, company_akzo, company_philips, company_van_houten, company_unilever)
df
```

### Task#6: Submit the project on Github

```{r}
write.csv(df, file = "refine_clean.csv")
```


