---
title: "Data Wrangling- Exercise#2"
author: "Danish Salam"
date: "July 7, 2017"
output: html_document
---
## Data Wrangling- Exercise#2

This document contains the tasks which are requested in exercise#2 in chapter#3.

### Load the data in RStudio
Save the data set as a CSV file called titanic_original.csv and load it in R Studio into a data frame.

```{r message = FALSE}
#load library
library(dplyr)
library(tidyr)

#load refine csv and hold it in local data frame.
org_df = read.csv("titanic_original.csv")
df <- tbl_df(org_df)
```

### Task #1: Port of embarkation

The embarked column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S. (Caution: Sometimes a missing value might be read into R as a blank or empty string.)

```{r df}
df <- df %>% 
  mutate(embarked = gsub("^$", "S", embarked))

#unique company name list
df %>% select(embarked) %>% unique()
```

### Task #2: Age

Calculate the mean of the Age column and use that value to populate the missing values.

```{r}
df <- df %>% 
  mutate(age = if_else(is.na(age), mean(df$age, na.rm=TRUE), age))
df %>% select(age) %>% unique()
```

### Task #3: Lifeboat
Fill these empty slots with a dummy value e.g. the string 'NA'.

```{r}
df <- df %>% 
  mutate(boat = gsub("^$", "NA", boat))

#unique boat list
df %>% select(boat) %>% unique()

```

### Task #4: Cabin
Create a new column has_cabin_number which has 1 if there is a cabin number, and 0 otherwise.

```{r}
df <- df %>% 
  mutate(as_cabin_number = if_else(grepl("^$", cabin), 0, 1))

df %>% 
  select(as_cabin_number) %>% distinct
```

### Task #6: Submit the project on Github
Include your code, the original data as a CSV file titanic_original.csv, and the cleaned up data as a CSV file called titanic_clean.csv.

```{r}
write.csv(df, file = "titanic_clean.csv")
```


