---
title: "Chapter-3 Exercise-1"
author: "Danish Salam"
date: "July 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling- Exercise#1

This document contains the tasks which are requested in exercise#1 in chapter#3.

### Load the data in RStudio
Save the data set as a CSV file called refine_original.csv and load it in R Studio into a data frame.

```{r message = FALSE}
#load library
library(dplyr)
library(tidyr)

#load refine csv and hold it in local data frame.
org_df = read.csv("refine_original.csv")
df <- tbl_df(org_df)
```


### Task # 1: Clean up brand names

Clean up the 'company' column so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).

```{r df}
df$company <-gsub("\\w+ps$", "philips", df$company, ignore.case = TRUE)
df$company <-gsub("^ak\\s*\\w+", "akzo", df$company, ignore.case = TRUE)
df$company <-gsub("^van\\s\\w+", "van houten", df$company, ignore.case = TRUE)
df$company <-gsub("^un\\w+", "unilever", df$company, ignore.case = TRUE)

#unique company name list
df %>% select(company) %>% unique()
```

### Task#2: Separate product code and number

Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively

```{r}
#split the code number column into two columns prodcut code and number respectively.
df <- df %>%
  separate(Product.code...number, c("ProductCode", "ProductNumber"), "-")
df %>% select(ProductCode, ProductNumber) %>% unique()
```

### Task#3: Add product categories
You learn that the product codes actually represent the following product categories:

    p = Smartphone
    v = TV
    x = Laptop
    q = Tablet
    
In order to make the data more readable, add a column with the product category for each record.

```{r}
#add a new column product category. Product category will be identified by this product code.
df <- df %>% 
  mutate(category = case_when(
    .$ProductCode=="p" ~ "Smartphone",
    .$ProductCode=="v" ~ "TV",
    .$ProductCode=="x" ~ "Laptop",
    .$ProductCode=="q" ~ "Tablet"
  ))

#unique prodcut category list
df %>% select(category) %>% unique()

```

### Task#4: Add full address for geocoding
You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

```{r}

df <-df %>% 
  mutate(full_address = paste(address, city, country, sep = ', '))
df %>% 
  select(full_address) %>% distinct
```

### Task#5: Create dummy variables for company and product category
Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.,

    Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever.

    Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet.


```{r}
df <- df %>% 
  mutate(company_philips = if_else(company=="philips", 1, 0)) %>% 
  mutate(company_akzo = if_else(company=="akzo", 1, 0)) %>% 
  mutate(company_van_houten = if_else(company=="van houten", 1, 0)) %>% 
  mutate(company_unilever = if_else(company=="unilever", 1, 0)) %>% 
  mutate(product_smartphone = if_else(ProductCode=="p", 1, 0)) %>% 
  mutate(product_tv = if_else(ProductCode=="v", 1, 0)) %>% 
  mutate(product_laptop = if_else(ProductCode=="x", 1, 0)) %>% 
  mutate(product_tablet = if_else(ProductCode=="q", 1, 0)) %>% 
  mutate_at(vars(matches("company_|product_")),funs(as.logical))
  
df
```

### Task#6: Submit the project on Github
nclude your code, the original data as a CSV file refine_original.csv, and the cleaned up data as a CSV file called refine_clean.csv.

```{r}
write.csv(df, file = "refine_clean.csv")
```


